Iterative Monomorphisation

### 2025-05-12 iterative_monomorphisation.tex
move appendix etc.
mroe

This example illustrates how an infinite number of new formulae can be generated from a simple initial problem.
%Although the example may seem contrived, similar problems frequently arise in practice. For example, the \sym{concat} function of Isabelle/HOL, which is characterised in the base case by the equation \(\sym{concat}\langle\alpha\rangle\; \sym{Nil}\langle\ty{list}(\alpha)\rangle = \sym{Nil}\langle\alpha\rangle\), exhibits the same behavior.
/Any reasonable implementation must limit the number of new type arguments, substitutions and formulae.

@@

\section{Appendix}

\begin{examplex}
This example runs an example through the detailed version of the iterative algorithm works. Consider the following initial problem:
\begin{itemize}
   \item[\(\langle1\rangle\)] \(\sym{f}\langle \ty{int}, \ty{nat}\rangle(-1, 3) \land \sym{f}\langle \ty{int}, \ty{int}\rangle(-1, -2)\)
   \item[\(\langle 2\rangle\)] \(\forall x: \alpha{,}\; y:\ty{list}(\alpha){,}\; z: \beta{,}\; \sym{f}\langle \ty{list}(\alpha), \alpha\rangle(y, x) \land \sym{f}\langle \alpha, \alpha \rangle(x, x) \land \sym{f}\langle \alpha, \beta\rangle(x, z)\)
\end{itemize}
%
\(M_{\text{new}}\) is initialised with \(\{f\mapsto (\ty{int}, \ty{nat}); f\mapsto (\ty{int}, \ty{int})\}\) and \(N_{\text{new}}\) with \(\{\langle 2\rangle\mapsto \{f\mapsto (\ty{list}(\alpha), \alpha); f\mapsto (\alpha, \alpha); f\mapsto (\alpha, \beta)\}\}\). Then we enter the main loop (assuming \Loop is at least 1). We iterate over each formula and call \FStep for each of them. Nothing happens for formula \(\langle1\rangle\) because it contains no type variables. For formula \(\langle2\rangle\), \(M_{\text{new}}\) and \(N_{\text{new}}(\langle2\rangle)\) are passed as argument to \SubstGen.

>>>>

This example illustrates how an infinite number of new formulae can be generated from a simple initial problem.
%Although the example may seem contrived, similar problems frequently arise in practice. For example, the \sym{concat} function of Isabelle/HOL, which is characterised in the base case by the equation \(\sym{concat}\langle\alpha\rangle\; \sym{Nil}\langle\ty{list}(\alpha)\rangle = \sym{Nil}\langle\alpha\rangle\), exhibits the same behavior.
Any reasonable implementation must limit the number of new type arguments, substitutions and formulae.
A more elaborate example is provided in Appendix~\ref{app}.%
\footnote{The appendix example will be included here in the camera-ready version if
the program chairs give use an extra page or two, as is customary.}

@@

\appendix

\section{Detailed Example}
\label{app}

\begin{examplex}\rm
This example runs an example through the detailed version of the iterative algorithm works. Consider the following initial problem:
\begin{itemize}
   \item[\(\langle1\rangle\)] \(\sym{f}\langle \ty{int}, \ty{nat}\rangle(-1, 3) \land \sym{f}\langle \ty{int}, \ty{int}\rangle(-1, -2)\)
   \item[\(\langle 2\rangle\)] \(\forall x: \alpha{,}\; y:\ty{list}(\alpha){,}\; z: \beta{,}\; \sym{f}\langle \ty{list}(\alpha), \alpha\rangle(y, x) \land \sym{f}\langle \alpha, \alpha \rangle(x, x) \land \sym{f}\langle \alpha, \beta\rangle(x, z)\)
\end{itemize}
%
\(M_{\text{new}}\) is initialised with \(\{f\mapsto (\ty{int}, \ty{nat}); f\mapsto (\ty{int}, \ty{int})\}\) and \(N_{\text{new}}\) with \(\{\langle 2\rangle\mapsto \{f\mapsto (\ty{list}(\alpha), \alpha); f\mapsto (\alpha, \alpha); f\mapsto (\alpha, \beta)\}\}\). Then we enter the main loop (assuming \Loop is at least 1). We iterate over each formula and call \FStep for each of them. Nothing happens for formula \(\langle1\rangle\) because it contains no type variables. For formula \(\langle2\rangle\), \(M_{\text{new}}\) and \(N_{\text{new}}(\langle2\rangle)\) are passed as argument to \SubstGen.
